Large and Infinitary Quotient Inductive-Inductive Types

### 2019-12-31 draft/paper.tex
paper

\newcommand{\bCon}{\boldsymbol{\Con}}
\newcommand{\bTy}{\boldsymbol{\Ty}}
\newcommand{\bSub}{\boldsymbol{\Sub}}
\newcommand{\bTm}{\boldsymbol{\Tm}}
\newcommand{\bGamma}{\boldsymbol{\Gamma}}
\newcommand{\bDelta}{\boldsymbol{\Delta}}
\newcommand{\bsigma}{\boldsymbol{\sigma}}
\newcommand{\bdelta}{\boldsymbol{\delta}}
\newcommand{\bepsilon}{\boldsymbol{\epsilon}}
\newcommand{\bt}{\boldsymbol{t}}
\newcommand{\bu}{\boldsymbol{u}}
\newcommand{\bA}{\boldsymbol{A}}
\newcommand{\ba}{\boldsymbol{a}}
\newcommand{\bb}{\boldsymbol{b}}
\newcommand{\bB}{\boldsymbol{B}}
\newcommand{\bid}{\boldsymbol{\id}}
\newcommand{\bemptycon}{\boldsymbol{\emptycon}}
\newcommand{\bSet}{\boldsymbol{\Set}}
\newcommand{\bU}{\boldsymbol{\U}}
\newcommand{\bEl}{\boldsymbol{\El}}
\newcommand{\bPii}{\boldsymbol{\Pi}}
\newcommand{\bPie}{\boldsymbol{\Pie}}
\newcommand{\bPiinf}{\boldsymbol{\Piinf}}
\newcommand{\bappi}{\boldsymbol{\mathsf{app}}}
\newcommand{\blami}{\boldsymbol{\mathsf{lam}}}

@@

\todo{P and Q in cwf, level lifts, think about revising cETT or compressing}

@@

For $A : \Set_i$ and $\bB : A \ra \bTy\,\bGamma$, we define $\bPie\,A\,\bB \boldsymbol{:} \bTy\,\bGamma$ as
the evident $A$-indexed direct product of $\bB$.

@@

For $A : \Set_i$ and $\bb : A \ra \bTm\,\bGamma\,\bU$, we aim to define
$\bPiinf\,A\,\bB \boldsymbol{:} \bTm\,\bGamma\,\bU$.

>>>>

\newcommand{\bCon}{\boldsymbol{\Con}}
\newcommand{\bTy}{\boldsymbol{\Ty}}
\newcommand{\bSub}{\boldsymbol{\Sub}}
\newcommand{\bTm}{\boldsymbol{\Tm}}
\newcommand{\bGamma}{\boldsymbol{\Gamma}}
\newcommand{\bDelta}{\boldsymbol{\Delta}}
\newcommand{\bsigma}{\boldsymbol{\sigma}}
\newcommand{\bdelta}{\boldsymbol{\delta}}
\newcommand{\bepsilon}{\boldsymbol{\epsilon}}
\newcommand{\bt}{\boldsymbol{t}}
\newcommand{\bu}{\boldsymbol{u}}
\newcommand{\bA}{\boldsymbol{A}}
\newcommand{\ba}{\boldsymbol{a}}
\newcommand{\bb}{\boldsymbol{b}}
\newcommand{\bB}{\boldsymbol{B}}
\newcommand{\bid}{\boldsymbol{\id}}
\newcommand{\bemptycon}{\boldsymbol{\emptycon}}
\newcommand{\bSet}{\boldsymbol{\Set}}
\newcommand{\bU}{\boldsymbol{\U}}
\newcommand{\bEl}{\boldsymbol{\El}}
\newcommand{\bPii}{\boldsymbol{\Pi}}
\newcommand{\bPie}{\boldsymbol{\Pie}}
\newcommand{\bPiinf}{\boldsymbol{\Piinf}}
\newcommand{\bappi}{\boldsymbol{\mathsf{app}}}
\newcommand{\blami}{\boldsymbol{\mathsf{lam}}}
\newcommand{\bId}{\boldsymbol{\Id}}

@@

\todo{revise cETT (compress), more mention of interp of term constructors in model, include missing iso-cleaving to type subst}

@@

For $A : \Set_i$ and $\bB : A \ra \bTy\,\bGamma$, we define $\bPie\,A\,\bB
\boldsymbol{:} \bTy\,\bGamma$ as the $A$-indexed direct product of $\bB$. Since
the indexing is given by a metatheoretic function, every component is given in the
evident pointwise way.

@@

For $A : \Set_i$ and $\bb : A \ra \bTm\,\bGamma\,\bU$, we aim to define
$\bPiinf\,A\,\bb \boldsymbol{:} \bTm\,\bGamma\,\bU$. The underlying functions
are:
\begin{alignat*}{3}
  & (\bPiinf\,A\,\bb)\,\ulGamma    &&:\equiv (a : A)\ra \bb\,a\,\ulGamma\\
  & (\bPiinf\,A\,\bb)\,\ulsigma    &&:\equiv \lambda\,a.\, \bb\,a\,\ulsigma\\
  & (\bPiinf\,A\,\bb)\,\ulA\       &&:\equiv \lambda\,\Gamma.\,(a : A)\ra \bb\,a\,\ulA\,(\Gamma\, a)\\
  & (\bPiinf\,A\,\bb)\,\ult        &&:\equiv \lambda\,a.\, \bb\,a\,\ult
\end{alignat*}
The preservation morphisms are as follows. Note that $\emptycon_{\bU} = \top$ and $\ext_{\bU}$ is
metatheoretic $\Sigma$.
\begin{alignat*}{3}
  &(\bPiinf\,A\,\bb)_{\emptycon}^{-1} && : \top\ra (\bPiinf\,A\,\bb)\,\emptycon\\
  %% &                               && : \top\ra (a : A) \ra \bb\,a\,\emptycon\\
  &(\bPiinf\,A\,\bb)_{\emptycon}^{-1} && :\equiv \lambda\,\_\,a.\,(\bb\,a)_{\emptycon}^{-1}\,\tt\\
  &(\bPiinf\,A\,\bb)_{\ext}^{-1} && : (\Gamma : (\bPiinf\,A\,\bb)\,\ulGamma)\times((\bPiinf\,A\,\bb)\,\ulA\,\Gamma)\\
  & && \hspace{0.5em}\ra (\bPiinf\,A\,\bb)\,(\ulGamma \ext \ulA)\\
  & (\bPiinf\,A\,\bb)_{\ext}^{-1} && :\equiv \lambda\,(\Gamma,A)\,a.\,(\bb\,a)_{\ext}^{-1}(\Gamma\,a,\,A\,a)
\end{alignat*}

The preservation of $\emptycon$ and $\blank\ext\blank$ here is in fact the main
point of divergence from \cite{kaposi2019constructing}. In ibid., substitutions
and terms are modeled as strict morphisms and types as displayed CwFs (with no
iso-cleaving). However, it is not the case that $(\bPiinf\,A\,\bb)\,\emptycon =
\top$, which is the statement of strict $\emptycon$-preservation. The left side
reduces to $(a : A)\ra \bb\,a\,\emptycon$, which is isomorphic to $\top$ but not
strictly equal to it. Likewise for $\ext$-preservation.

Hence, we are forced to
interpret terms as weak sections, which in turn forces us to interpret
types as isofibrations, since type substitution requires iso-cleaving.

\subsection{Identity}
For $\bt$ and $\bu$ in $\bTm\,\bGamma\,(\bEl\,\ba)$, we define $\bId\,\bt\,\bu
\boldsymbol{:} \bTm\,\bGamma\,\bU$ as expressing pointwise equality of weak
sections.
\begin{alignat*}{3}
& (\bId\,\bt\,\bu)\,\ulGamma &&:\equiv \bt\,\ulGamma = \bu\,\ulGamma\\
& (\bId\,\bt\,\bu)\,\ulA     && :\equiv \lambda\,e.\, \bt\,\ulA = \bu\,\ulA
\end{alignat*}
Above, $\bt\,\ulA = \bu\,\ulA$ is well-typed because of $e :
\bt\,\ulGamma = \bu\,\ulGamma$. For substitutions, we have to complete a square
of equalities:
\begin{alignat*}{3}
  (\bId\,\bt\,\bu)\,(\ulsigma : \Sub\,\ulGamma\,\ulDelta) : \bt\,\ulGamma = \bt\,\ulGamma \ra
       \bt\,\ulDelta = \bt\,\ulDelta
\end{alignat*}
This can be given by $\bt\,\ulsigma : \ba\,\ulsigma\,(\bt\,\ulGamma) =
\bt\,\ulDelta$ and $\bu\,\ulsigma : \ba\,\ulsigma\,(\bu\,\ulGamma) =
\bu\,\ulDelta$. The action on terms is analogous. We omit preservation morphisms
here as they are straightforward. Like $\bPiinf$, $\bId$ also does not support
strict preservation of $\emptycon$ and $\ext$. Equality reflection and $\boldsymbol{\refl :}
\bId\,\bt\,\bt$ are also evident.

\section{Model Theory of the Theory of Signatures}
